---
title: Measuring the effect of political actions to contrast the contagion of COVID19
author: "Angela Andreella, Michele Lambardi, Silvia De Nicolo"
date: "June 3, 2020"
output:
  ioslides_presentation:
    incremental: yes
    theme: cosmo
    transition: slower
    logo: unipd_logo.png
    widescreen: true
    smaller: true
    css: style.css
---

```{r setup, include=FALSE}
rm(list = ls())
path <- "~/GitHub/Statistical_Consulting/" #Insert your path
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
source(paste0(path, "Angela/packages.R"))
```

# Explorative Analysis {data-background=covid.jpg data-background-size=cover}

## Objective {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

Policymakers in different Countries have introduced <span style="color:rgb(17,57,219);">different political action</span> to contrast the COVID19 contagion.

1. What are the different **containment efforts** and is there a **strategies** resemblance across countries?

2. What is the **effect** of these policies on the contagion from a **global perspective**?

3. Has the **same action** lead to different results in the case of different **regions of Italy**?

## Data {data-background=back.jpg data-background-size=cover .flexbox .vcenter}


* COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at **Johns Hopkins University** for contagion data,

*  **Oxford** COVID-19 Government Response Tracker (OxCGRT) for policies tracking. 


## Containment strategies and resembling patterns {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

*  Dimension Reduction via <span style="color:rgb(17,57,219);">Polychoric PCA</span> for $11$ ordinal variables (from $0$ to $2$ or $0$ to $3$) indicating the stringency level of policies such as

     + **School**, **workplace** and **transport** closing and **event** cancellation;
  
     + **Gathering**, **stay-home** and internal/international **movement** restrictions;
  
     + **Information**, **Testing** and Contact **Tracing** campaigns. 


*  <span style="color:rgb(17,57,219);">Functional Data Co-Clustering</span> of the countries aligned to the first contagion (from the 10th day before contagion).

## Containment strategies {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, echo=FALSE}
load(paste0(path,"Presentation/Data/cluster20paesi.Rdata"))
load(paste0(path,"Presentation/Data/SAVEPLOTS.Rdata"))
```

## Containment strategies {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

**Restriction-based** policies on one hand, **Tracing** and **Testing** policies on the other hand.

```{r, echo=FALSE}
ggplot(dpca, aes(x=PC, y=Loadings, fill=policies, group=position))+
 geom_bar(stat="identity", position="dodge")+
  coord_flip()+scale_fill_manual("", breaks=c(apol, rpol), 
                                 values=c(gray[1], peach[1], gray[2],                                           peach[2], gray[3:6], peach[3], gray[7:8]))
```
 
## Resembling patterns {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, echo=FALSE}
load(paste0(path,"Presentation/Data/Silvia_maps.RData"))
grid.arrange(na2, europe2,                  
             arrangeGrob(kor2, sin2, nrow = 2, ncol=1),
             ncol = 3, nrow=1, widths=c(3,4,2)) 
```


## Resembling patterns {data-background=back.jpg data-background-size=cover .flexbox .vcenter}
```{r, fig.align='center', echo=FALSE}
ggarrange(restrict, testt, trace, ncol=3, nrow=1, common.legend = TRUE, legend="bottom")
```

# Worldwide Analysis {data-background=covid.jpg data-background-size=cover}

## Motivation {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, echo=FALSE, warning = FALSE, comment = FALSE, silent = T}
load(paste0(path, "Presentation/Data/out.RData"))
a <- ggplot(dat, aes(x = date, y = active_count/pop, group = Clusters, color = Clusters)) + geom_smooth(method = 'loess',formula = 'y ~ x') + scale_x_date(date_labels = "%b %d") + theme_classic() + ylab("Active/population size") + xlab("Date") 

ggplotly(a)
```


## Model {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

<center>
<span style="color:rgb(17,57,219);">GENERALIZED POISSON MIXED MODEL for Overdispersed Count Data</span>
</div> 
</center>

**WHY?**

- We analyze the number of **Active person**, i.e., Confirmed - Deaths - Recovered $\rightarrow$ **Count Dependent Variable** (Generalized Poisson Model);


- The data are observed for each **country nested within clusters** during $103$ **days** $\rightarrow$ **Mixed Model**.


- **Confounders** from the World Bank Open Data and Oxford Data:
  - **FIXED**: 
    - Population density;
  - **LONGITUDINAL**: 
    - Economic: Income Support and Debt/contract relief for households;
    - Health: Emergency Investment in healthcare and Investment in vaccines.


## Which policies acted better? {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r,echo=FALSE, warning = FALSE, comment = FALSE, silent = T}
out0 <- as.data.frame(ggeffect(mod1, terms = "workplace_closingF"))
out0$Value <- c("No measures", "Recommend closing or work from home", "require closing for some sector" ,"Require closing all-but-essential workplaces")
out1 <-as.data.frame(ggeffect(mod1, terms = "gatherings_restrictionsF"))
out1$Value <- c("No measures", "> 1000", "100-1000 people", "10-100 people", "< 10 people")
out2 <- as.data.frame(ggeffect(mod1, terms = "transport_closingF"))[c(1,3),]
out2$Value <- c("No measures", "Require closing")
out3 <- as.data.frame(ggeffect(mod1, terms = "stay_home_restrictionsF"))[c(1,4),]
out3$Value <- c("No measures", "Minimal exceptions")
out4 <- as.data.frame(ggeffect(mod1, terms = "testing_policyF"))
out4$Value <- c("No measures", "Specific criteria", "Symptoms", "Open")
out5 <- as.data.frame(ggeffect(mod1, terms = "contact_tracingF"))
out5$Value <- c("No measures", "Limited", "Comprehensive")

OUT <- rbind(out0, out1, out2, out3, out4, out5)
OUT$Policies <- c(rep("Workplace Closing", nrow(out0)),
                  rep("No Gathering", nrow(out1)),
                  rep("No Transport", nrow(out2)),  
                  rep("Stay Home", nrow(out3)),
                  rep("Testing", nrow(out4)),
                  rep("Tracing", nrow(out5)))

colnames(OUT)[1] <- "Strength"

#OUT$cols <- c("slateblue1", "slateblue2", "slateblue3", "slateblue4")

ggplot(OUT, aes(Policies, predicted, fill = Strength)) +
  geom_bar(stat = "summary", fun = "mean", 
           position = position_dodge(width = 0.9)) +
  scale_y_continuous(name = "Predicted counts of Active People after 14 days", limits = c(0, 1700))+ geom_errorbar(stat = "summary", fun.data = "mean_sdl", 
                fun.args = list(mult = 2),
                position =  position_dodge(width =1)) +
  geom_text(aes(label=Value, group = Strength),     hjust = -0.1, size = 3,
            position = position_dodge(width = 1),
            inherit.aes = TRUE)  + coord_flip() + scale_fill_viridis(discrete=T) +
  theme_minimal() + ylab("") 
```

## Who acted promptly? {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r echo=FALSE, warning = FALSE, comment = FALSE, silent = T}
load(paste0(path, "Presentation/Data/plot_map.RData"))
grid.arrange(na2, europe2,                 
             arrangeGrob(kor2, sin2, nrow = 2, ncol=1),
             ncol = 3, nrow=1, widths=c(3,4,2))    
```

## Who acted promptly? {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, echo=FALSE, warning = FALSE, comment = FALSE, silent = T}
load(paste0(path, "Presentation/Data/plot_mod.RData"))

subplot(ggplotly(p1, tooltip = c("colour", "y")) %>% layout(showlegend = FALSE, title = ""),
 ggplotly(p2, tooltip = c("colour", "y"))%>% layout(showlegend = FALSE,title = ""), 
ggplotly(p3, tooltip = c("colour", "y")) %>% layout(showlegend = FALSE, title = "Stay Home           Testing            Tracing"), shareY = T
) 
```

## Take home message {data-background=back.jpg data-background-size=cover .flexbox .vcenter}



- <span style="color:rgb(17,57,219);">Lockdown policies work!</span> respect to impose no measure;

- Strong Testing and Tracing policies lead to discovering more infected people (luckily!)



- **Korea and Singapore are the best countries that acted properly**;

- Sweden, Germany, Portugal, and Greece better than the other UE countries;

- The USA, and Canada better than the other UE countries except for Sweden and Germany.


# Italy Case {data-background=covid.jpg data-background-size=cover}

## Introduction {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

<div class="blue2">Background</div>


* Italian **regions**, ethernal **divide**

* Lockdown almost simultaneous, excepted the **Red Zone**

* First cases in **Lombardia** and **Lazio** hubs

<div class="blue2">Problems</div>


* Policies have **no variability** between regions

* **Baseline control**: some regions start from worse situations

* Cannot estimates some effects as for the nations case

* To our defense, integration between databases came lately

* Instrumental variables, more correct but tricky approach


## Approach {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

* **Phase "1" versus Phase "0"** comparison


* **Auto-regression**: modelling active cases given past numbers
   * Related but not quite to the R0 index
   
* **Random effects** for region and date, standard panel approach


* Assuming policies effects seen **~14 days later**


* Controlling for **testing** frequency

```{r, warning=FALSE, echo = FALSE, include = FALSE}
load(paste0(path, "Presentation/Data/finaleregioni.RData"))
load(paste0(path, "Presentation/Data/maps.RData"))
map <- map %>% filter(country=="Italy")

npaz <- 30

dat <- dat %>%
   mutate(
      monthspoint=14/30*log(npaz/active)/log(ratepoint),
      monthslower=14/30*log(npaz/active)/log(ratelower),
      monthsupper=14/30*log(npaz/active)/log(rateupper)
   )

last <- dat %>%
   filter(date==max(date)) %>%
   left_join(map, by="region") %>%
   st_as_sf()
```

## Speed of contagion {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, warning=FALSE, echo = FALSE}
last %>%
   arrange(ratepoint) %>%
   mutate(region=factor(region, levels=region)) %>%
   ggplot(aes(x=ratepoint, y=region)) +
   geom_point(aes(x=ratepoint)) +
   geom_errorbar(aes(xmin=ratelower, xmax=rateupper, group=region)) +
   geom_text(aes(x=ratelower, label=paste(region, " ")), hjust="right") +
   theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white"),
        panel.grid.major.x = element_line(color="grey"),
        panel.grid.minor.x = element_line(color="lightgrey"),
        # panel.grid.major.y = element_line(color="grey"),
        legend.position = "none") +
   xlab("active cases in two weeks, expected") +
   scale_x_continuous(limits = c(0,1), labels = function(x) paste0(ifelse(x<1,"","+"), (x-1)*100, "%")) # Multiply by 100 
```

## ETV, Estimated Time to Victory {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, warning=FALSE, echo = FALSE}
last %>%
   arrange(monthspoint) %>%
   mutate(region=factor(region, levels=region)) %>%
   ggplot(aes(x=monthspoint, y=region)) +
   geom_point(aes(x=monthspoint)) +
   geom_errorbar(aes(xmin=monthslower, xmax=monthsupper, group=region)) +
   geom_text(aes(x=monthsupper, label=paste(" ", region)), hjust="left") +
   theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white"),
        panel.grid.major.x = element_line(color="grey"),
        panel.grid.minor.x = element_line(color="lightgrey"),
        # panel.grid.major.y = element_line(color="grey"),
        legend.position = "none") +
   xlab(paste0("months to ", npaz, "-or-less active cases, expected, approximate")) +
   xlim(0, 24)
```

## Joint view {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, warning=FALSE, echo = FALSE}
npaz <- 30
last %>%
   ggplot(aes(x=ratepoint, y=monthspoint)) +
   geom_point() +
   geom_errorbar(aes(ymin=monthslower, ymax=monthsupper, group=region)) +
   # geom_errorbar(aes(xmin=ratelower, xmax=rateupper, group=region)) +
   # geom_text(aes(x=monthsupper, label=paste(" ", region)), hjust="left") +
   theme(
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white"),
        panel.grid.major.x = element_line(color="grey"),
        panel.grid.minor.x = element_line(color="lightgrey"),
        panel.grid.major.y = element_line(color="grey"),
        panel.grid.minor.y = element_line(color="lightgrey"),
        # panel.grid.major.y = element_line(color="grey"),
        legend.position = "none") +
   xlab(paste0("expected cases in two weeks")) +
   ylab(paste0("months to ", npaz, "-or-less active cases")) +
   ylim(0, 20) +
   scale_x_continuous(limits = c(0.25,1), labels = function(x) paste0(ifelse(x<1,"","+"), (x-1)*100, "%")) +
   geom_text(aes(label=ifelse(ratepoint<0.7,"",paste0("  ", region))), hjust="left")
   #+
   # xlim(0, 24)
```

## A map of criticality {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, warning=FALSE, echo = FALSE}
# $ geometry           :sfc_MULTIPOLYGON of length 51; first list element: List of 1
#   ..$ :List of 1
#   .. ..$ : num [1:245, 1:2] -1111066 -1110710 -1096162 -1095175 -1012095 ...
#   ..- attr(*, "class")= chr  "XY" "MULTIPOLYGON" "sfg"
#  - attr(*, "sf_column")= chr "geometry"
#  - attr(*, "relation_to_geometry")= Factor w/ 3 levels "field","lattice",..: NA NA NA NA NA NA NA NA NA NA ...
#   ..- attr(*, "names")= chr  "geo_id" "fips_state" "name" "lsad" ...
ggplotly(
   last %>%
      mutate(value=(ratepoint-1)*100) %>%
      as_Spatial() %>%
      st_as_sf() %>%
      ggplot() +
      geom_sf(aes(fill=value), color="black", lwd=0.5) +
      scale_fill_gradientn(colors = c("white", "darkblue"), labels = function(x) paste0(ifelse(x<0,"","+"), x, "%")) +
      theme(axis.text.x = element_blank(),
            # axis.title.x = element_blank(),
            axis.title.y = element_blank(),
           axis.text.y = element_blank(),
           axis.ticks = element_blank(),
           rect = element_blank(),
           legend.position = "none") +
      # geom_sf_text(aes(label=ifelse(region=="Basilicata", "?", "")), color="white") +
      xlab(paste0("active cases expected in two weeks"))
)
```

## Anything weird? {data-background=back.jpg data-background-size=cover .flexbox .vcenter}

```{r, warning=FALSE, echo = FALSE}
cutpoint <- "2020-03-22"
ggplotly(
   dat %>%
      mutate(value=100*(ratepoint-1)) %>%
      filter(date >= cutpoint) %>%
      ggplot(aes(x=date, y=value, color=region)) +
      geom_line() +
      theme(legend.position = "none") +
      scale_x_date(date_labels = "%b %d") +
      scale_y_continuous(limits = c(-100,100*max(-1+dat$ratepoint[dat$date >= cutpoint])), labels = function(x) paste0(ifelse(x<0,"","+"), x, "%")) +
      xlab("") +
      ylab("active cases in two weeks, expected")
   )
   #+
   #  +
   # geom_sf_text(aes(label=ifelse(region=="Basilicata", "?", ""))) +
   # xlab(paste0("active cases expected in two weeks"))
```

# Conclusions {data-background=covid.jpg data-background-size=cover}
