---
title: Measuring the effect of different political actions to contrast the contagion of COVID19
author: "Angela Andreella, Michele Lambardi, Silvia De Nicolo"
date: "June 3, 2020"
output:
  ioslides_presentation:
    incremental: yes
    theme: cosmo
    transition: slower
  slidy_presentation:
    incremental: yes
  beamer_presentation:
    incremental: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

Policy makers in different Countries have introduced different political action to contrast the COVID19 contagion.

1. What are the different containment efforts and is there a strategies resemblance across countries?

2. What is the effect of these policies on the contagion from a global perspective?

3. Has the same action lead to different results in the case of different regions of Italy?


## Data

* COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at **Johns Hopkins University** for contagion data,

*  **Oxford** COVID-19 Government Response Tracker (OxCGRT) for policies tracking. 


## 1. Containment strategies and resembling patterns

*  Dimension Reduction via Polychoric PCA for 11 ordinal variables (from 0 to 2 or 0 to 3) indicating the stringency level of policies such as

     + School, workplace and transport closing and event cancellation;
  
     + Gathering, stay-home and internal/international movement restrictions;
  
     + Information, Testing and Contact Tracing campaigns. 


*  Function Data Co-Clustering of the countries aligned to the first contagion (from the 10th day before contagion).

## Containment strategies

```{r, include=F}
rm(list = ls())
require(COVID19, quietly=T)
require(ggplot2, quietly=T)
require(dplyr, quietly=T)
require(countrycode, quietly=T)
require(maps, quietly=T)
require(rnaturalearth, quietly=T)
require(rnaturalearthdata, quietly=T)
require(ggpubr, quietly=T)
require(gridExtra, quietly=T)
require(psych, quietly=T)
require(viridis, quietly=T)

load("cluster20paesi.Rdata")
load("/Users/siviadenicolo/Documents/GitHub/Statistical_Consulting/Silvia/SAVEPLOTS.Rdata")

```

## Containment strategies

```{r,  echo=FALSE}
ggplot(dpca, aes(x=PC, y=Loadings, fill=policies, group=position))+
 geom_bar(stat="identity", position="dodge")+
  coord_flip()+scale_fill_manual("", breaks=c(apol, rpol), 
                                 values=c(gray[1], peach[1], gray[2],                                           peach[2], gray[3:6], peach[3], gray[7:8]))

```

Co-clustering of policies variables: Restriction- based policies on one hand, Tracing and Testing policies on the other hand.

## Resembling patterns


```{r, echo=FALSE}
grid.arrange(na2, europe2,                  
             arrangeGrob(kor2, sin2, nrow = 2, ncol=1),
             ncol = 3, nrow=1, widths=c(3,4,2)) 

```


## Resembling patterns
```{r, fig.align='center', echo=FALSE, fig.width = 8}
ggarrange(restrict, testt, trace, ncol=3, nrow=1, common.legend = TRUE, legend="bottom")

```
