---
title: "Italy case"
author: ''
date: ''
output:
  beamer_presentation:
    incremental: yes
  ioslides_presentation:
    incremental: yes
    theme: cosmo
    transition: slower
  slidy_presentation:
    incremental: yes
---

# Italy Case

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

## Background

* Italian regions, ethernal divide
* Lockdown almost simultaneous, excepted the Red Zone
* First cases in Lombardia and Lazio hubs

## Problem

* Policies have no variability between regions
* Baseline control: some regions start from worse situations
* Cannot estimates some effects as for the nations case
* To our defense, integration between databases came lately
* Instrumental variables, more correct but tricky approach

## Approach

* Phase "1" versus Phase "0" comparison
* Auto-regression: modelling active cases given past numbers
   * Related but not quite to the R0 index
* Random effects for region and date, standard panel approach
* Assuming policies effects seen ~14 days later
* Controlling for testing frequency

```{r, include=F}
rm(list = ls())
# devtools::install_github('ropensci/plotly')
require(plotly, quietly = T)
require(sp, quietly = T)
require(dplyr, quietly = T)
require(glmmTMB, quietly = T)
require(COVID19, quietly = T)
require(effects, quietly = T)
require(ggplot2, quietly = T)
require(sf, quietly = T)
require(sfheaders)
require(rnaturalearth, quietly = T)
require(rnaturalearthdata, quietly = T)
load("../Michele/finaleregioni.RData")
load("../Michele/maps.RData")
map <- map %>% filter(country=="Italy")

npaz <- 30

dat <- dat %>%
   mutate(
      monthspoint=14/30*log(npaz/active)/log(ratepoint),
      monthslower=14/30*log(npaz/active)/log(ratelower),
      monthsupper=14/30*log(npaz/active)/log(rateupper)
   )

last <- dat %>%
   filter(date==max(date)) %>%
   left_join(map, by="region") %>%
   st_as_sf()
```

## Speed of contagion

```{r, warning=F}
last %>%
   arrange(ratepoint) %>%
   mutate(region=factor(region, levels=region)) %>%
   ggplot(aes(x=ratepoint, y=region)) +
   geom_point(aes(x=ratepoint)) +
   geom_errorbar(aes(xmin=ratelower, xmax=rateupper, group=region)) +
   geom_text(aes(x=ratelower, label=paste(region, " ")), hjust="right") +
   theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white"),
        panel.grid.major.x = element_line(color="grey"),
        panel.grid.minor.x = element_line(color="lightgrey"),
        # panel.grid.major.y = element_line(color="grey"),
        legend.position = "none") +
   xlab("active cases in two weeks, expected") +
   scale_x_continuous(limits = c(0,1), labels = function(x) paste0(ifelse(x<1,"","+"), (x-1)*100, "%")) # Multiply by 100 
```

## ETV, Estimated Time to Victory

```{r, warning=F}
last %>%
   arrange(monthspoint) %>%
   mutate(region=factor(region, levels=region)) %>%
   ggplot(aes(x=monthspoint, y=region)) +
   geom_point(aes(x=monthspoint)) +
   geom_errorbar(aes(xmin=monthslower, xmax=monthsupper, group=region)) +
   geom_text(aes(x=monthsupper, label=paste(" ", region)), hjust="left") +
   theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white"),
        panel.grid.major.x = element_line(color="grey"),
        panel.grid.minor.x = element_line(color="lightgrey"),
        # panel.grid.major.y = element_line(color="grey"),
        legend.position = "none") +
   xlab(paste0("months to ", npaz, "-or-less active cases, expected, approximate")) +
   xlim(0, 24)
```

## Joint view

```{r, warning=F}
npaz <- 30
last %>%
   ggplot(aes(x=ratepoint, y=monthspoint)) +
   geom_point() +
   geom_errorbar(aes(ymin=monthslower, ymax=monthsupper, group=region)) +
   # geom_errorbar(aes(xmin=ratelower, xmax=rateupper, group=region)) +
   # geom_text(aes(x=monthsupper, label=paste(" ", region)), hjust="left") +
   theme(
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white"),
        panel.grid.major.x = element_line(color="grey"),
        panel.grid.minor.x = element_line(color="lightgrey"),
        panel.grid.major.y = element_line(color="grey"),
        panel.grid.minor.y = element_line(color="lightgrey"),
        # panel.grid.major.y = element_line(color="grey"),
        legend.position = "none") +
   xlab(paste0("expected cases in two weeks")) +
   ylab(paste0("months to ", npaz, "-or-less active cases")) +
   ylim(0, 20) +
   scale_x_continuous(limits = c(0.25,1), labels = function(x) paste0(ifelse(x<1,"","+"), (x-1)*100, "%")) +
   geom_text(aes(label=ifelse(ratepoint<0.7,"",paste0("  ", region))), hjust="left")
   #+
   # xlim(0, 24)
```

## A map of criticality

```{r, warning=F}
# $ geometry           :sfc_MULTIPOLYGON of length 51; first list element: List of 1
#   ..$ :List of 1
#   .. ..$ : num [1:245, 1:2] -1111066 -1110710 -1096162 -1095175 -1012095 ...
#   ..- attr(*, "class")= chr  "XY" "MULTIPOLYGON" "sfg"
#  - attr(*, "sf_column")= chr "geometry"
#  - attr(*, "relation_to_geometry")= Factor w/ 3 levels "field","lattice",..: NA NA NA NA NA NA NA NA NA NA ...
#   ..- attr(*, "names")= chr  "geo_id" "fips_state" "name" "lsad" ...
ggplotly(
   last %>%
      mutate(value=(ratepoint-1)*100) %>%
      as_Spatial() %>%
      st_as_sf() %>%
      ggplot() +
      geom_sf(aes(fill=value), color="black", lwd=0.5) +
      scale_fill_gradientn(colors = c("white", "darkblue"), labels = function(x) paste0(ifelse(x<0,"","+"), x, "%")) +
      theme(axis.text.x = element_blank(),
            # axis.title.x = element_blank(),
            axis.title.y = element_blank(),
           axis.text.y = element_blank(),
           axis.ticks = element_blank(),
           rect = element_blank(),
           legend.position = "none") +
      # geom_sf_text(aes(label=ifelse(region=="Basilicata", "?", "")), color="white") +
      xlab(paste0("active cases expected in two weeks"))
)
```

## Anything weird?

```{r, warning=F}
cutpoint <- "2020-03-22"
ggplotly(
   dat %>%
      mutate(value=100*(ratepoint-1)) %>%
      filter(date >= cutpoint) %>%
      ggplot(aes(x=date, y=value, color=region)) +
      geom_line() +
      theme(legend.position = "none") +
      scale_x_date(date_labels = "%b %d") +
      scale_y_continuous(limits = c(-100,100*max(-1+dat$ratepoint[dat$date >= cutpoint])), labels = function(x) paste0(ifelse(x<0,"","+"), x, "%")) +
      xlab("") +
      ylab("active cases in two weeks, expected")
   )
   #+
   #  +
   # geom_sf_text(aes(label=ifelse(region=="Basilicata", "?", ""))) +
   # xlab(paste0("active cases expected in two weeks"))
```
