---
title: "Second Report"
author: "Angela Andreella"
date: "21/05/2020"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
#load package
source("Angela/packages.R")

#load variables
load("Angela/Data/var.RData")

#load data
load("Data/db_with_R0.RData")

dat <- df
#add Cluster Silvia aligned

uno <- c("KOR", "SGP")
due <- c("DEU", "SWE")
tre <- c("CAN", "GRC", "PRT","USA")
quattro <- c("ESP", "GBR", "IRL", "ITA", "NLD")
cinque <- c("AUT", "BEL", "CHE", "DNK", "FIN", "FRA", "NOR")
states_to_sel <- c(uno,due,tre,quattro,cinque)

datA <- dat %>% filter(id %in% states_to_sel)

datA$Clusters <- ifelse(datA$id %in% uno, "Cl1", 
                        ifelse(datA$id %in% due, "Cl2", 
                               ifelse(datA$id %in% tre, "Cl3", 
                                      ifelse(datA$id %in% quattro, "Cl4", 
                                             "Cl5"))))
dat <- datA

#Some steps 
dat$Clusters <- factor(dat$Clusters)

dat_shape <- dat %>%
  left_join(dat %>%
              filter(confirmed >= 1) %>%
              group_by(id) %>%
              summarise(date.start=min(date)),
            by="id") %>%
  mutate(date2=date-date.start) %>%
  as.data.frame()

dat_shape <- add_R0_ML(dat_shape)

dat_shape <- 
  dat_shape %>% arrange(id, date) %>%
  group_by(id) %>%
  mutate(confirmed_lag = Hmisc::Lag(confirmed, shift  = -14))

dat_shape <- 
  dat_shape %>% 
#  group_by(id) %>%
  mutate(confirmed_prop = confirmed_lag/pop,
         active = pmax(confirmed - deaths - recovered, 0))

dat_shape$confirmed_count <- ave(dat_shape$confirmed, dat_shape$id, FUN=function(x) pmax(0, c(0, diff(x))))

dat_shape <- 
  dat_shape %>% arrange(id, date) %>%
  group_by(id) %>%
  mutate(confirmed_count_lag = Hmisc::Lag(confirmed_count, shift  = -14),
         active_lage = Hmisc::Lag(active, shift  = -14))

```

# Motivation

So, the aim is to understand how the lockdown policies influences the contagions. We consider the aligned data respect to the first confirmed case, we have the following situation:

```{r}
a <- ggplot(dat_shape, aes(x = date, y = confirmed_count, group = Clusters, color = Clusters)) + geom_smooth() + scale_x_date(date_labels = "%b %d") + theme_classic() + ylab("Date") + xlab("Confirmed")

a <- direct.label(a, list(last.qp,cex=1,dl.trans(x=x+0.1)))

#ggplotly(a)
```

```{r}
a <- ggplot(dat_shape, aes(x = date, y = active, group = Clusters, color = Clusters)) + geom_smooth() + scale_x_date(date_labels = "%b %d") + theme_classic() + ylab("Date") + xlab("Confirmed")

a <- direct.label(a, list(last.qp,cex=1,dl.trans(x=x+0.1)))

#ggplotly(a)
```

Also, we lag the number of confirmed respect to $14$ days, in order to consider the influences of the restrictions imposed at time $t$ on number of confirmed at time $t+14$, in order to make a correct impact.

```{r}
pca_EC <- polychoric(dat_shape[,var_EC[1:2]])
matPCA_ec <- cbind(rep(0,2),pca_EC$tau)
dat_shape$ox.E1_Income.support_f <- as.factor(dat_shape$ox.E1_Income.support)
dat_shape$ox.E1_Income.support_f <- recode_factor(dat_shape$ox.E1_Income.support_f,
                                                  "0" = paste0(matPCA_ec[1,1]),
                                                  "1" = paste0(matPCA_ec[1,2]),
                                                  "2" = paste0(matPCA_ec[1,3]))

dat_shape$ox.E1_Income.support_f <- as.numeric(dat_shape$ox.E1_Income.support_f)
dat_shape$ox.E2_Debt.contract.relief_f <- as.factor(dat_shape$ox.E2_Debt.contract.relief)
dat_shape$ox.E2_Debt.contract.relief_f <- recode_factor(dat_shape$ox.E2_Debt.contract.relief_f,
                                                        "0" = paste0(matPCA_ec[2,1]),
                                                        "1" = paste0(matPCA_ec[2,2]),
                                                        "2" = paste0(matPCA_ec[2,3]))
dat_shape$ox.E2_Debt.contract.relief_f <- as.numeric(dat_shape$ox.E2_Debt.contract.relief_f)
dat_shape$ox.E3_Fiscal.measures_log <- log(dat_shape$ox.E3_Fiscal.measures +1)
dat_shape$ox.E4_International.support <- ifelse(is.na(dat$ox.E4_International.support), 0, dat_shape$ox.E4_International.support)
dat_shape$ox.E4_International.support_log <- log(dat_shape$ox.E4_International.support +1)
dat_shape$pop_log <- log(dat_shape$pop +1)
pca_age <- princomp(na.omit(dat_shape[,var_FIX[2:3]]), cor = TRUE)
dat_shape$pca_age <- predict(pca_age,newdata = dat_shape[,var_FIX[2:3]])[,1]
dat_shape$pop_density_log <- log(dat_shape$pop_density + 1)
dat_shape$pop_death_rate_1000 <- dat_shape$pop_death_rate * 1000
dat_shape$gdp_log <- log(dat_shape$gdp + 1)
pca_hs <- princomp(na.omit(dat_shape[,var_HS]), cor = TRUE)
dat_shape$pca_hs <- predict(pca_hs,newdata = dat_shape[,var_HS])[,1]
dat_shape$school_closingF <- as.factor(dat_shape$school_closing)
dat_shape$cancel_eventsF <- as.factor(dat_shape$cancel_events)
dat_shape$gatherings_restrictionsF <- as.factor(dat_shape$gatherings_restrictions)
dat_shape$transport_closingF <- as.factor(dat_shape$transport_closing)
dat_shape$stay_home_restrictionsF <- as.factor(dat_shape$stay_home_restrictions)
dat_shape$workplace_closingF <- as.factor(dat_shape$workplace_closing)
dat_shape$internal_movement_restrictionsF <- as.factor(dat_shape$internal_movement_restrictions)
dat_shape$testing_policyF <- as.factor(dat_shape$testing_policy)
dat_shape$contact_tracingF <- as.factor(dat_shape$contact_tracing)

#pca policies

policies <- c("school_closing", "workplace_closing", "cancel_events", "gatherings_restrictions","transport_closing", "stay_home_restrictions", "internal_movement_restrictions")
W <- polychoric(dat_shape[,policies])
W$tau
p3 <- principal(r = W$rho, nfactors = 6) 
p3$loadings #prime 3 componenti
cbind(policies, p3$loadings)

scores <- as.matrix(dat_shape[,policies]) %*% p3$weights
dat_shape$pca_LD <- scores[,1]

#pca economics
policiesEC <- c("ox.E1_Income.support", "ox.E2_Debt.contract.relief")

W <- polychoric(dat_shape[,policiesEC])
W$tau
p3 <- principal(r = W$rho, nfactors = 1) 
p3$loadings #prime 3 componenti
cbind(policies, p3$loadings)

scores <- as.matrix(dat_shape[,policiesEC]) %*% p3$weights
dat_shape$pca_EC <- scores[,1]
```




```{r}
var_EC <- c("ox.E1_Income.support_f","ox.E2_Debt.contract.relief_f", var_EC[5:6], "ox.E3_Fiscal.measures_log", "ox.E4_International.support_log")

var_FIX <- c("pca_age", "pop_log", "pop_density_log", "hosp_beds", "pop_death_rate_1000", "gdp_log", "health_exp")

var_HS <- "pca_hs"

var_LD <- c("school_closingF","workplace_closingF","cancel_eventsF",
            "gatherings_restrictionsF","transport_closingF","stay_home_restrictionsF",
            "internal_movement_restrictionsF","testing_policyF","contact_tracingF")
```

So, we plot the correlation matrix of the variables that we will consider for the model:

```{r}
cormat <- round(cor(dat_shape[,c(var_EC, "confirmed_prop", var_FIX, var_HS, "pca_EC", "pca_LD")],use = "na.or.complete"),6)
dat_cor <- melt(cormat)

p <- ggplot(data = dat_cor, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

ggplotly(p)

```


```{r}
ggplot(dat_shape, aes(confirmed_count, group = Clusters)) + geom_histogram() 
ggplot(dat_shape, aes(active, group = Clusters)) + geom_histogram() 

```


# Model

```{r}
ICCest(date2, confirmed_prop, dat_shape) # 0.3293131
ICCest(id, confirmed_prop, dat_shape) #0.1389142
ICCest(Clusters, confirmed_prop, dat_shape) #0.06034613
```

Plot to understand the variability respect date

```{r}
abc<-aggregate(confirmed_count ~date2, dat_shape, mean)
bdata <- dat_shape
bdata <- merge(bdata,abc,by="date2")
bdata$colorBox <- ifelse(bdata$confirmed_count.y>= mean(na.omit(bdata$confirmed_count.x)), "#56B4E9", "#009E73")

bdata$date2 <- as.factor(bdata$date2)
ggplot(bdata, aes(x=date2, y=confirmed_count.x, fill = colorBox)) +
    geom_boxplot() +
  scale_color_manual(values=rainbow(6))+ theme(legend.position="none")
```
and respect Clusters:

```{r}
abc<-aggregate(confirmed_prop ~Clusters, dat_shape, mean)
bdata <- dat_shape
bdata <- merge(bdata,abc,by="Clusters")
bdata$colorBox <- ifelse(bdata$confirmed_prop.y>= mean(na.omit(bdata$confirmed_prop.x)), "#56B4E9", "#009E73")

bdata$Clusters <- as.factor(bdata$Clusters)
ggplot(bdata, aes(x=Clusters, y=confirmed_prop.x, fill = colorBox)) +
    geom_boxplot() +
  scale_color_manual(values=rainbow(6))+ theme(legend.position="none")
```

and id:

```{r}
abc<-aggregate(confirmed_prop ~id, dat_shape, mean)
bdata <- dat_shape
bdata <- merge(bdata,abc,by="id")
bdata$colorBox <- ifelse(bdata$confirmed_prop.y>= mean(na.omit(bdata$confirmed_prop.x)), "#56B4E9", "#009E73")

bdata$id <- as.factor(bdata$id)
ggplot(bdata, aes(x=id, y=confirmed_prop.x, fill = colorBox)) +
    geom_boxplot() +
  scale_color_manual(values=rainbow(6))+ theme(legend.position="none")
```

## Model with Clusters from functional Clustering Silvia

```{r}
ggplot(dat_shape, aes(confirmed_count, fill = Clusters)) +
  geom_histogram() +
  scale_x_log10() +
  facet_grid(Clusters ~ ., margins=TRUE, scales="free_y")
```

Zero-Inflated Mixed Models


```{r}
ggplot(dat_shape, aes(confirmed_count, fill = Clusters)) +
  geom_histogram() +
  scale_x_log10() +
  facet_grid(Clusters ~ ., margins=TRUE, scales="free_y")
```

Per confrontare i due modelli con stessa parte random usiamo LRT, se invece anche cambia quella random cAIC.  si segue The Top-Down Strategy from LINEAR MIXED MODELS by Brady T. West et al. Prima analizza la parte della media e poi quella della covarianza quindi prima aggiungiamo tutto fisso

```{r}

f <- as.formula(paste("confirmed_count_lag", "~", 
                      paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(1,3,6)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters"))

mod1 <- glmmTMB(f, dat_shape, family="poisson")
check_overdispersion(mod1)
mod1 <- glmmTMB(f, dat_shape, family="nbinom2")
check_collinearity(mod1)
summary(mod1)

f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(1,3,6)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters"))
mod1 <- glmmTMB(f, dat_shape, family="nbinom2")
check_collinearity(mod1)
summary(mod1)

f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(1)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters"))
mod1 <- glmmTMB(f, dat_shape, family="nbinom2")
check_collinearity(mod1)
summary(mod1)

f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(1)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(1,3:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters"))
mod13 <- glmmTMB(f, dat_shape, family="nbinom2")
check_collinearity(mod13)
summary(mod13)
check_autocorrelation(mod13)

f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(1)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(1,3:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters + (1|id) + (1|date2) + (1|Clusters)"))
mod14 <- glmmTMB(f, dat_shape, family="nbinom2")
check_collinearity(mod14)
summary(mod14)
check_autocorrelation(mod14)
check_singularity(mod14)
model_performance(mod14)

f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(1)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(1,3:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters + (1|id) + (1|date2)"))
mod15 <- glmmTMB(f, dat_shape, family="nbinom2")
check_collinearity(mod15)
summary(mod15)
check_autocorrelation(mod15)
check_singularity(mod15)
model_performance(mod15)


f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                   #  "+", paste(c(var_FIX[c(1)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(1,3:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters + (1|id) + (1|date2)"))
mod16 <- glmmTMB(f, dat_shape, family="nbinom2")
check_collinearity(mod16)
summary(mod16)
check_autocorrelation(mod16)
check_singularity(mod16)
model_performance(mod16)


f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                   #  "+", paste(c(var_FIX[c(1)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(1,3:4,7:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters + (1|id) + (1|date2)"))
mod16 <- glmmTMB(f, dat_shape, family="nbinom2")

f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                   #  "+", paste(c(var_FIX[c(1)]), collapse=" + "),
                      "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(1,3:4,7:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters + (1+pca_EC|id) + (1|date2)"))
mod17 <- glmmTMB(f, dat_shape, family="nbinom2")

f <- as.formula(paste("confirmed_count_lag", "~", 
                   #   paste(c(var_EC[c(3)]), collapse=" + "), 
                 #    "+", paste(c(var_FIX[c(3)]), collapse=" + "),
                    #  "+", paste(c(var_HS[1]), collapse=" + "),
                  #  "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(6,4,7:9)]),collapse=" + "),
                      "+ offset(log(pop))  + Clusters + pca_EC + (1|id/Clusters) + (1+log(mkt_close+1) + pca_EC|date2)"))
mod18 <- glmmTMB(f, dat_shape, family="nbinom2")

check_collinearity(mod18)
summary(mod18)
check_autocorrelation(mod18)
check_singularity(mod18)
model_performance(mod18)
compare_performance(mod16,mod17,mod18)
g1 <- glht_glmmTMB(mod16, linfct = mcp(Clusters = "Tukey"))
summary(g1)

f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(3,6)]), collapse=" + "),
                     # "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(6,4,5,8:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters + (1|id/Clusters) + (1|date2)")) 

mod18 <- glmmTMB(f, dat_shape, family="nbinom2")
summary(mod18)

```


```{r}

f <- as.formula(paste("confirmed_count_lag", "~", 
                     # paste(c(var_EC[c(3)]), collapse=" + "), 
                   #  "+", paste(c(var_FIX[c(3)]), collapse=" + "),
                     # "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(4, 6:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters + (1+pca_EC|id/Clusters) + (1 + mkt_close|date2)"))

mod <- glmmTMB(f, dat_shape, family="nbinom2")
summary(mod)
mod1 <- glmmTMB(f, dat_shape, family="poisson")
summary(mod1)
mod2 <- glmmTMB(f, dat_shape, family="poisson", ziformula = ~.1)
summary(mod2)
mod3 <- glmmTMB(f, dat_shape, family="nbinom2", ziformula = ~.1)

summary(zi_mod)
AICtab(zi_mod)
check_overdispersion(mod1)


```




```{r}
f <- as.formula(paste("log(confirmed_lag + 1)", "~", 
                      paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(6,7)]), collapse=" + "),
                      "+", paste(c(var_HS), collapse=" + "),
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ offset(log(pop)) + (1|id/Clusters) + (1|date2)"))
model <- glmer(f, data = dat_shape,family = quasipoisson,na.action = na.omit)
summary(model)
rsq(model,adj=TRUE)
```


Using active?

```{r}
f <- as.formula(paste("active_lage", "~", 
                   #   paste(c(var_EC[c(3)]), collapse=" + "), 
                  #   "+", paste(c(var_FIX[c(2,3)]), collapse=" + "),
                  #    "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(1,4, 5,6:9)]),collapse=" + "),
                      "+ Clusters  + pop_density_log + pca_EC + (0 + pca_LD|id) + (1|date2) + (1| Clusters)"))

mod1 <- glmmTMB(f, dat_shape, family="nbinom2", offset = log(active + 1))

save(mod1, file = "final_model_grazie_dio.RData")

summary(mod1)

check_collinearity(mod1)
check_overdispersion(mod1)
a <- ggpredict(mod1, "Clusters")
check_autocorrelation(mod1)
check_collinearity(mod1)
check_zeroinflation(mod1)

plot(a,ci = F)

dat_shape$pred <- predict(mod1, newdata = dat_shape)

a<-ggplot(dat_shape) + geom_line(aes(x = date2, y = pred/active, group = id, color = Clusters))
ggplotly(a)


g1 <- glht_glmmTMB(mod1, linfct = mcp(Clusters = "Tukey"))
summary(g1)
```



con PCA?

```{r}




dat_shape %>% ggplot(aes(x=date2, y=pca_LD, group=Clusters, color=Clusters)) + geom_smooth()

```

```{r}


dat_shape %>% ggplot(aes(x=date2, y=pca_EC, group=Clusters, color=Clusters)) + geom_smooth()

```

QUESTO!!!

```{r}

dat_shape$obs_effect <- 1:nrow(dat_shape)
dat_shape$ox.E1_Income.support_f <- as.factor(dat_shape$ox.E1_Income.support_f)
dat_shape$ox.E2_Debt.contract.relief_f <- as.factor(dat_shape$ox.E2_Debt.contract.relief_f)


f <- as.formula(paste("confirmed_count_lag", "~", 
                    #  paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(3,6)]), collapse=" + "),
                     # "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(6,4,5,8:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters + (1|id/Clusters) + (1|date2) + (1|obs_effect)"))
uno <- c("ITA", "FRA", "ESP", "BEL")
due <- c("KOR" ,"SGP")
tre <- c( "DEU" ,"SWE", "NLD", "FIN","GBR")
quattro <- c("CAN", "USA")
dat_shape1 <- dat_shape %>% filter(id %in% c(uno,due,tre,quattro))
dat_shape1$Clusters1 <- ifelse(dat_shape1$id %in% uno, "Cl1",
                               ifelse(dat_shape1$id %in% due, "Cl2",
                                      ifelse(dat_shape1$id %in% tre,"Cl3", "Cl4")))
f <- as.formula(paste("confirmed_count_lag", "~", 
                     # paste(c(var_EC[c(3)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(6)]), collapse=" + "),
                     # "+", paste(c(var_HS[1]), collapse=" + "),
                   # "+ log(ox.H4_Emergency.investment.in.healthcare +1)",
                     "+", paste(c(var_LD[c(6,8:9)]),collapse=" + "),
                      "+ offset(log(pop)) + pca_EC + Clusters1 + (1|id/Clusters1) + (1|date2) + (1|obs_effect)"))

model <- glmmTMB(f, data = dat_shape1,family = poisson(link = "log"), 
                 na.action = na.omit)
summary(model)
check_overdispersion(model)
MuMIn::r.squaredGLMM(model,adj=TRUE)
plot_model(model)
pr<-ggeffect(model, "Clusters", type = "re")

plot(pr,ci = FALSE,colors = "hero")

pr<-ggeffect(model, "gatherings_restrictionsF")

plot(pr,ci = TRUE,colors = "hero")

pr<-ggeffect(model, "testing_policyF")

plot(pr,ci = TRUE,colors = "hero")

pr<-ggeffect(model, "contact_tracingF")

plot(pr,ci = TRUE,colors = "hero")


g1 <- glht_glmmTMB(model, linfct = mcp(Clusters1 = "Tukey"))
summary(g1)

```

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf") 
world <- world %>% filter(iso_a3 %in% states_to_sel)
world$id <- world$iso_a3 

dat_shape$confFIT <- ifelse(dat_shape$Clusters == "Cl1", pr$predicted[1],
                    ifelse(dat_shape$Clusters == "Cl2", pr$predicted[2],
                           ifelse(dat_shape$Clusters == "Cl3", pr$predicted[3],
                                  ifelse(dat_shape$Clusters == "Cl4", pr$predicted[4], pr$predicted[5]))))

datMap <- left_join(world, dat_shape, by="id")

ggplot(data = datMap) + geom_sf(aes(fill = confFIT))
```


PROVE DI MERDA QUI:


#Clusters 1 Korea Singapore

```{r}
dat1 <- dat_shape %>% filter(id %in% uno)
f <- as.formula(paste("confirmed_count_lag", "~", 
                     # paste(c(var_EC[c(3)]), collapse=" + "), 
                     #"+", paste(c(var_FIX[6]), collapse=" + "),
                    #  "+", paste(c(var_HS), collapse=" + "),
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ pca_LD + offset(log(pop))"))
model <- glm(f, data = dat1,family = poisson(link = "log"), na.action = na.omit)
summary(model)
MuMIn::r.squaredGLMM(model,adj=TRUE)
plot_model(model)
```
#Clusters 2 Korea "DEU" "SWE"

```{r}
dat2 <- dat_shape %>% filter(id %in% due)
f <- as.formula(paste("confirmed_count_lag", "~", 
                     # paste(c(var_EC[c(3)]), collapse=" + "), 
                     #"+", paste(c(var_FIX[6]), collapse=" + "),
                    #  "+", paste(c(var_HS), collapse=" + "),
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ offset(log(pop)) + (1|date)"))
model <- glmer(f, data = dat1,family = poisson(link = "log"), na.action = na.omit)
summary(model)
MuMIn::r.squaredGLMM(model,adj=TRUE)
plot_model(model)
```

#Clusters 3 Korea "CAN"  "GRC"  "PRT"  " USA"

```{r}
dat3 <- dat_shape %>% filter(id %in% tre)
f <- as.formula(paste("confirmed_count_lag", "~", 
                     # paste(c(var_EC[c(3)]), collapse=" + "), 
                     #"+", paste(c(var_FIX[6]), collapse=" + "),
                    #  "+", paste(c(var_HS), collapse=" + "),
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ offset(log(pop)) + (1|date)"))
model <- glmer(f, data = dat1,family = poisson(link = "log"), na.action = na.omit)
summary(model)
MuMIn::r.squaredGLMM(model,adj=TRUE)
plot_model(model)
```

```{r}
comparison <- c("Cl1 - Cl2 = 0",
                "Cl1 - Cl3 = 0",
                "Cl1 - Cl4 = 0",
                "Cl1 - Cl5 = 0",
                "Cl2 - Cl3 = 0",
                "Cl2 - Cl4 = 0",
                "Cl2 - Cl5 = 0",
               "Cl3 - Cl4 = 0",
                "Cl3 - Cl5 = 0",
                "Cl4 - Cl5 = 0")
pairs(emmeans::emmeans(model, "Clusters"))

test<-multcomp::glht(model, linfct = mcp(Clusters = comparison),adjusted="holm")
summary(test)
```




```{r}
f <- as.formula(paste("log(confirmed_lag + 1)", "~", 
                     # paste(c(var_EC[c(3)]), collapse=" + "), 
                    # "+", paste(c(var_FIX[c(6,7)]), collapse=" + "),
                    #  "+", paste(c(var_HS), collapse=" + "),
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ offset(log(pop)) + Clusters"))
fm <- mixed_model(fixed = f, random = ~ 1 | id/Clusters + 1|date2 , data = dat_shape,
                  family = poisson())
summary(fm)
rsq(fm,adj=TRUE)
```

```{r}
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}


f <- as.formula(paste("log(confirmed_lag + 1)", "~", 
                   #   paste(c(var_EC[c(3)]), collapse=" + "), 
                   #  "+", paste(c(var_FIX[c(6,7)]), collapse=" + "),
                   #   "+", paste(c(var_HS), collapse=" + "),
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ offset(log(pop)) + Clusters + (1|id/Clusters) + (1|date2)"))
model <- glmer(f, data = dat_shape,family = poisson(link = "log"),na.action = na.omit)
overdisp_fun(model)
summary(model)
rsq(model,adj=TRUE)
```

```{r}
f <- as.formula(paste("log(confirmed_lag+1)", "~", paste(c(var_EC), collapse=" + "), 
                     "+", paste(c(var_FIX), collapse=" + "),
                      "+", paste(c(var_HS), collapse=" + "),
                 #     "+", "offset(log(pop))", 
                     "+", paste(c(var_LD),collapse=" + "),
                      "+ date2 + Clusters"))

model <- glm(f, data = dat_shape,family = binomial(link = "logit"), offset = log(pop))
summary(model)
rsq(model,adj=TRUE)
#0.9725014
plot(model)
```

```{r}
f <- as.formula(paste("confirmed_prop", "~", paste(c(var_EC[c(1:4)]), collapse=" + "), 
                     "+", paste(c(var_FIX), collapse=" + "),
                   #   "+", paste(c(var_HS), collapse=" + "),
                   #   "+", "Clusters",
                     "+", paste(c(var_LD), collapse=" + "),
                      "+  offset(confirmed/pop) + Clusters"))
model <- glm(f, data = dat_shape,family = quasibinomial(link = "logit"))
summary(model)
rsq(model,adj=TRUE)
```

```{r}
comparison <- c("Cl1 - Cl2 = 0",
                "Cl1 - Cl3 = 0",
                "Cl1 - Cl4 = 0",
                "Cl1 - Cl5 = 0",
                "Cl2 - Cl3 = 0",
                "Cl2 - Cl4 = 0",
                "Cl2 - Cl5 = 0",
                "Cl3 - Cl4 = 0",
                "Cl3 - Cl5 = 0",
                "Cl4 - Cl5 = 0")
test<-multcomp::glht(model, linfct = mcp(Clusters = comparison),adjusted="holm")
summary(test)
# > uno
# [1] KOR SGP
# > due
# [1] DEU SWE
# > tre
# [1] CAN GRC PRT USA
# > quattro
# [1] ESP GBR IRL ITA NLD
# > cinque
# [1] AUT BEL CHE DNK FIN FRA NOR
```

```{r}
dat_shape$testing_policyF <- recode_factor(dat_shape$testing_policyF,
                                                  "0" = "A",
                                                  "1" = "B",
                                                  "2" = "C",
                                                  "3" = "D")
model <- glm(f, data = dat_shape,family = quasibinomial(link = "logit"))
test<-multcomp::glht(model, linfct = mcp(testing_policyF = comparison),adjusted="holm")
summary(test)
```


## Model separati per Cluster Scarpa

```{r}
uno <- c("ITA", "FRA")
due <- c("KOR" ,"SGP")
tre <- c( "DEU" ,"SWE")
dat_shapeScarpa <- dat_shape %>% filter(id %in% c(uno,due,tre)) 
dat_shapeScarpa1 <- dat_shape %>% filter(id %in% uno)
dat_shapeScarpa2 <- dat_shape %>% filter(id %in% due)
dat_shapeScarpa3 <- dat_shape %>% filter(id %in% tre)
dat_shapeScarpa$Clusters1 <- 

f <- as.formula(paste("log(confirmed)", "~", paste(c(var_EC[c(1:3)]), collapse=" + "), 
                    # "+", paste(c(var_FIX), collapse=" + "),
                   #   "+", paste(c(var_HS), collapse=" + "),
                   #   "+", "Clusters",
                     "+", paste(c(var_LD), collapse=" + "),
                      "+  offset(log(pop)) + Clusters1"))

model <- glm(f, data = dat_shapeScarpa,family = quasibinomial(link = "logit"))
summary(model)
rsq(model)
#           R2m       R2c
#[1,] 0.2796572 0.7935903
plot(model)
```

```{r}
comparison <- c("A - B = 0",
                "A - C = 0",
                "B - C = 0")
test<-multcomp::glht(model, linfct = mcp(testing_policyF = comparison),adjusted="holm")
summary(test)
```

```{r}

f <- as.formula(paste("confirmed_prop", "~", paste(c(var_EC[c(1,4)]), collapse=" + "), 
                     "+", paste(c(var_FIX[7]), collapse=" + "),
                      "+", paste(c(var_HS), collapse=" + "),
                   #   "+", "Clusters",
                     "+", paste(c(var_LD[c(1,2,3,5,6,7,8)]), collapse=" + "),
                      "+ date2 + id"))

dat_shapeScarpa1$testing_policyF <- recode_factor(dat_shapeScarpa1$testing_policyF,
                                                  "0" = "A",
                                                  "1" = "B",
                                                  "2" = "C",
                                                  "3" = "D")
model <- glm(f, data = dat_shapeScarpa2,family = quasibinomial(link = "logit"))
summary(model)
MuMIn::r.squaredGLMM(model)
#           R2m       R2c
#[1,] 0.2796572 0.7935903
plot(model)
```

```{r}
comparison <- c("A - B = 0",
                "A - C = 0",
                "B - C = 0")
test<-multcomp::glht(model, linfct = mcp(testing_policyF = comparison),adjusted="holm")
summary(test)
```
```{r}

f <- as.formula(paste("log(confirmed_lag)", "~", paste(c(var_EC[c(1,4)]), collapse=" + "), 
                     "+", paste(c(var_FIX[7]), collapse=" + "),
                      "+", paste(c(var_HS), collapse=" + "),
                   #   "+", "Clusters",
                     "+", paste(c(var_LD[c(1,2,3,5,6,7,8)]), collapse=" + "),
                      "+ date2 + id"))

dat_shapeScarpa1$testing_policyF <- recode_factor(dat_shapeScarpa1$testing_policyF,
                                                  "0" = "A",
                                                  "1" = "B",
                                                  "2" = "C",
                                                  "3" = "D")
model <- glm(f, data = dat_shapeScarpa3,family = quasibinomial(link = "logit"))
summary(model)
MuMIn::r.squaredGLMM(model)
#           R2m       R2c
#[1,] 0.2796572 0.7935903
plot(model)
```

```{r}
comparison <- c("A - B = 0",
                "A - C = 0",
                "B - C = 0")
test<-multcomp::glht(model, linfct = mcp(testing_policyF = comparison),adjusted="holm")
summary(test)
```

#Totale ma con Clusters 


```{r}
dat_shapeScarpa$Clusters <- ifelse(dat_shapeScarpa$id %in% uno, "Cl1",
                                   ifelse(dat_shapeScarpa$id %in% due, "Cl2","Cl3"))


dat_shapeScarpa$Clusters <- as.factor(dat_shapeScarpa$Clusters)
f <- as.formula(paste("confirmed_prop", "~", paste(c(var_EC[c(1,4)]), collapse=" + "), 
                     "+", paste(c(var_FIX[3]), collapse=" + "),
                   #  "+", paste(c(var_HS), collapse=" + "),
                   #   "+", "Clusters",
                     "+", paste(c(var_LD), collapse=" + "),
                      "+  offset(confirmed/pop) + Clusters"))

model <- glm(f, data = dat_shapeScarpa,family = quasibinomial(link = "logit"))
summary(model)
MuMIn::r.squaredGLMM(model)
#           R2m       R2c
#[1,] 0.2796572 0.7935903
plot(model)
```

```{r}
comparison <- c("Cl1 - Cl2 = 0",
                "Cl1 - Cl3 = 0",
                "Cl2 - Cl3 = 0")
test<-multcomp::glht(model, linfct = mcp(Clusters = comparison),adjusted="holm")
summary(test)
```


