---
title: "Report Stat_Consulting"
author: "Angela Andreella"
date: "16/05/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
#load package
source("Angela/packages.R")

#load variables
load("Angela/Data/var.RData")

#load data
load("Data/db_with_R0.RData")

dat <- df
#add Cluster Silvia aligned

uno <- c( "CAN", "DEU", "GBR", "SWE", "USA")
due <- c("IRL", "IRN", "ITA", "NLD", "TUR")
tre <- c( "AUT", "BEL", "CHE", "DNK", "ESP", "FIN", "FRA", "NOR","RUS")
quattro <- c("KOR", "SGP")
cinque <- c("GRC", "PRT", "ROU", "SAU")
states_to_sel <- c(uno,due,tre,quattro,cinque)

datA <- dat %>% filter(id %in% states_to_sel)

datA$Clusters <- ifelse(datA$id %in% uno, "Cl1", 
                        ifelse(datA$id %in% due, "Cl2", 
                               ifelse(datA$id %in% tre, "Cl3", 
                                      ifelse(datA$id %in% quattro, "Cl4", 
                                             "Cl5"))))
dat <- datA

#Some steps 
dat$Clusters <- factor(dat$Clusters)

dat_shape <- dat %>%
  left_join(dat %>%
              filter(confirmed >= 1) %>%
              group_by(id) %>%
              summarise(date.start=min(date)),
            by="id") %>%
  mutate(date2=date-date.start) %>%
  as.data.frame()

dat_shape <- add_R0_ML(dat_shape)

dat_shape <- 
  dat_shape %>% arrange(id, date) %>%
  group_by(id) %>%
  mutate(R0_lag = dplyr::lag(R0, n = 14, default = NA))

```

# Motivation

So, the aim is to understand how the lockdown policies influences the changing of the index $R0$, the reproduction number. We consider the aligned data respect to the first confirmed case, we have the following situation:

```{r}
a<-dat_shape %>% ggplot(aes(x=date, y=R0_lag, group=Clusters, color=Clusters)) + geom_smooth()

ggplotly(a)
```

Also, we lag the R0 respect to $14$ days, in order to consider the influences of the restrictions imposed at time $t$ on $R0$ at time $t+14$, in order to make a correct impact.

# Exploratory Analysis

So, we have the following variables to consider:

Economic ones (longitudinal):

```{r}
var_EC
```

Demographic/fixed ones:

```{r}
var_FIX
```

Health ones (longitudinal):

```{r}
var_HS
```

Lockdown ones:

```{r}
var_LD
```

plus the number of tests conducted, i.e., tests.

## Economic variables

```{r}
p <- dat_shape %>% ggplot(aes(x=date, y=ox.E1_Income.support, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
```

```{r}
p <- dat_shape %>% ggplot(aes(x=R0_lag, y=ox.E1_Income.support)) +
  geom_smooth() 

ggplotly(p)
```

```{r}
p <- dat_shape %>% ggplot(aes(x=date2, y=ox.E2_Debt.contract.relief, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
```

We will transform these two first economic variables as continous variables:

```{r}
pca_EC <- polychoric(dat_shape[,var_EC[1:2]])
matPCA_ec <- cbind(rep(0,2),pca_EC$tau)
dat_shape$ox.E1_Income.support_f <- as.factor(dat_shape$ox.E1_Income.support)
dat_shape$ox.E1_Income.support_f <- recode_factor(dat_shape$ox.E1_Income.support_f,
                                                  "0" = paste0(matPCA_ec[1,1]),
                                                  "1" = paste0(matPCA_ec[1,2]),
                                                  "2" = paste0(matPCA_ec[1,3]))

dat_shape$ox.E1_Income.support_f <- as.numeric(dat_shape$ox.E1_Income.support_f)
dat_shape$ox.E2_Debt.contract.relief_f <- as.factor(dat_shape$ox.E2_Debt.contract.relief)
dat_shape$ox.E2_Debt.contract.relief_f <- recode_factor(dat_shape$ox.E2_Debt.contract.relief_f,
                                                        "0" = paste0(matPCA_ec[2,1]),
                                                        "1" = paste0(matPCA_ec[2,2]),
                                                        "2" = paste0(matPCA_ec[2,3]))
dat_shape$ox.E2_Debt.contract.relief_f <- as.numeric(dat_shape$ox.E2_Debt.contract.relief_f)

```

```{r}
p <- dat_shape %>% ggplot(aes(x=date, y=ox.E2_Debt.contract.relief_f, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
```

```{r}
p <- dat_shape %>% ggplot(aes(x=date2, y=ox.E3_Fiscal.measures, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
#log scale
dat_shape$ox.E3_Fiscal.measures_log <- log(dat_shape$ox.E3_Fiscal.measures +1)
```

```{r}
p <- dat_shape %>% ggplot(aes(x=date2, y=ox.E4_International.support, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
#log scale and put 0 where NA 
dat_shape$ox.E4_International.support <- ifelse(is.na(dat$ox.E4_International.support), 0, dat_shape$ox.E4_International.support)
dat_shape$ox.E4_International.support_log <- log(dat_shape$ox.E4_International.support +1)
```


```{r}
p <- dat_shape %>% ggplot(aes(x=date2, y=mkt_close, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
```

```{r}
p <- dat_shape %>% ggplot(aes(x=date2, y=mkt_volume, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
```

## Fix variables

```{r}
p <- ggplot(dat_shape, aes(x=Clusters, y=log(pop), fill=Clusters)) + geom_boxplot() 

ggplotly(p)

#log
dat_shape$pop_log <- log(dat_shape$pop +1)

```


```{r}
p <- ggplot(dat_shape, aes(x=Clusters, y=pop_65, fill=Clusters)) + geom_boxplot() 


ggplotly(p)
```

```{r}
p <- ggplot(dat_shape, aes(x=Clusters, y=pop_age, fill=Clusters)) + geom_boxplot() 


ggplotly(p)
```

pca for these two *age* variables:

```{r}
pca_age <- princomp(na.omit(dat_shape[,var_FIX[2:3]]), cor = TRUE)
dat_shape$pca_age <- predict(pca_age,newdata = dat_shape[,var_FIX[2:3]])[,1]
```


```{r}
p <- ggplot(dat_shape, aes(x=Clusters, y=pop_density, fill=Clusters)) + geom_boxplot() 
#log
dat_shape$pop_density_log <- log(dat_shape$pop_density + 1)
ggplotly(p)
```

Togli Singapore. e metti log

```{r}
p <- ggplot(dat_shape, aes(x=Clusters, y=hosp_beds, fill=Clusters)) + geom_boxplot() 

ggplotly(p)

```

```{r}
p <- ggplot(dat_shape, aes(x=Clusters, y=pop_death_rate, fill=Clusters)) + geom_boxplot() 

ggplotly(p)
dat_shape$pop_death_rate_1000 <- dat_shape$pop_death_rate * 1000
```

```{r}
p <- ggplot(dat, aes(x=Clusters, y=gdp, fill=Clusters)) + geom_boxplot() 

ggplotly(p)

#log
dat_shape$gdp_log <- log(dat_shape$gdp + 1)
```


```{r}
p <- ggplot(dat_shape, aes(x=Clusters, y=health_exp, fill=Clusters)) + geom_boxplot() 


ggplotly(p)
```

```{r}
p <- ggplot(dat_shape, aes(x=Clusters, y=health_exp_oop, fill=Clusters)) + geom_boxplot() 

ggplotly(p)
```

## Health variables

```{r}
p <- dat_shape %>% ggplot(aes(x=date2, y=ox.H4_Emergency.investment.in.healthcare, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
```

```{r}
p <- dat_shape %>% ggplot(aes(x=date2, y=ox.H5_Investment.in.vaccines, group=Clusters, color=Clusters)) +
  geom_smooth() 

ggplotly(p)
```

pca:

```{r}
pca_hs <- princomp(na.omit(dat_shape[,var_HS]), cor = TRUE)
dat_shape$pca_hs <- predict(pca_hs,newdata = dat_shape[,var_HS])[,1]
```

## Lockdown variables

```{r}
dat_shape$school_closingF <- as.factor(dat_shape$school_closing)
dat_shape$cancel_eventsF <- as.factor(dat_shape$cancel_events)
dat_shape$gatherings_restrictionsF <- as.factor(dat_shape$gatherings_restrictions)
dat_shape$transport_closingF <- as.factor(dat_shape$transport_closing)
dat_shape$stay_home_restrictionsF <- as.factor(dat_shape$stay_home_restrictions)
dat_shape$workplace_closingF <- as.factor(dat_shape$workplace_closing)
dat_shape$internal_movement_restrictionsF <- as.factor(dat_shape$internal_movement_restrictions)
dat_shape$testing_policyF <- as.factor(dat_shape$testing_policy)
dat_shape$contact_tracingF <- as.factor(dat_shape$contact_tracing)
```

So, we consider these new set of variables:

```{r}
var_EC <- c("ox.E1_Income.support_f","ox.E2_Debt.contract.relief_f", var_EC[5:6], "ox.E3_Fiscal.measures_log", "ox.E4_International.support_log")

var_FIX <- c("pca_age", "pop_log", "pop_density_log", "hosp_beds", "pop_death_rate_1000", "gdp_log", "health_exp")

var_HS <- "pca_hs"

var_LD <- c("school_closingF","workplace_closingF","cancel_eventsF",
            "gatherings_restrictionsF","transport_closingF","stay_home_restrictionsF",
            "internal_movement_restrictionsF","testing_policyF","contact_tracingF")
```

So, we plot the correlation matrix of the variables that we will consider for the model:

```{r}
cormat <- round(cor(dat_shape[,c(var_EC, "R0_lag", var_FIX, var_HS, var_LD)],use = "na.or.complete"),6)
dat_cor <- melt(cormat)

p <- ggplot(data = dat_cor, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

ggplotly(p)

```


# Model

The data are observed for each country nested within date. 

- Two-level model: the units of analysis (Level 1), countries, are nested within clusters (Level 2), date;

- The variability of the data comes from nested sources;

- The Intraclass Correlation Coefficient (ICC) is equal to $0.3910876$ for date, equals $0.04668614$ for id and $ 0.02533497$ for Clusters.

```{r}
ICCest(date2, R0_lag, dat_shape) #0.3910876
ICCest(id, R0_lag, dat_shape) #0.04668614
ICCest(Clusters, R0_lag, dat_shape) #0.04668614
```

Plot to understand the variability respect date

```{r}
abc<-aggregate(R0 ~date2, dat_shape, mean)
bdata <- dat_shape
bdata <- merge(bdata,abc,by="date2")
bdata$colorBox <- ifelse(bdata$R0.y>= mean(na.omit(bdata$R0.x)), "#56B4E9", "#009E73")

bdata$date2 <- as.factor(bdata$date2)
ggplot(bdata, aes(x=date2, y=R0.x, fill = colorBox)) +
    geom_boxplot() +
  scale_color_manual(values=rainbow(6))+ theme(legend.position="none")
```

and respect Clusters:

```{r}
abc<-aggregate(R0 ~Clusters, dat_shape, mean)
bdata <- dat_shape
bdata <- merge(bdata,abc,by="Clusters")
bdata$colorBox <- ifelse(bdata$R0.y>= mean(na.omit(bdata$R0.x)), "#56B4E9", "#009E73")

bdata$Clusters <- as.factor(bdata$Clusters)
ggplot(bdata, aes(x=Clusters, y=R0.x, fill = colorBox)) +
    geom_boxplot() +
  scale_color_manual(values=rainbow(6))+ theme(legend.position="none")
```

and id:

```{r}
abc<-aggregate(R0 ~id, dat_shape, mean)
bdata <- dat_shape
bdata <- merge(bdata,abc,by="id")
bdata$colorBox <- ifelse(bdata$R0.y>= mean(na.omit(bdata$R0.x)), "#56B4E9", "#009E73")

bdata$id <- as.factor(bdata$id)
ggplot(bdata, aes(x=id, y=R0.x, fill = colorBox)) +
    geom_boxplot() +
  scale_color_manual(values=rainbow(6))+ theme(legend.position="none")
```

**How to choose the random and fixed part?**

The problem is much more complicated than in linear regression because selection on the covariance structure is not straightforward due to computational issues and boundary problems arising from positive semidefinite constraints on covariance matrices.

-Conditional AIC (Package cAIC4): The conditional AIC is also appropriate for choosing between a simple null model without any random effects and a complex model incorporating random effects,

-Boostrap (R Package pbkrtest): Model comparison of nested models using parametric bootstrap methods. Implemented for some commonly applied model types.

```{r}
#dat_shape$Clusters <- recode_factor(dat_shape$Clusters,
#                              "Cl1" = "testing_based",
#                              "Cl2" = "restrinctionsStrong_testingMild_based",
#                              "Cl3" = "mild_levels_all",
#                              "Cl4" = "testing_tracing_based",
#                              "Cl5" = "restriction_tracingMild")
```

```{r}
f <- as.formula(paste("R0_lag", "~", paste(c(var_EC[c(4)]), collapse=" + "), 
                     "+", paste(c(var_FIX[c(1:3)]), collapse=" + "),
                     # "+", paste(c(var_HS), collapse=" + "),
                   #   "+", "Clusters",
                     "+", paste(c(var_LD), collapse=" + "),
                      "+ (1 |date2) + (1|Clusters/id)"))

model <- lmer(f, data = dat_shape,REML=FALSE)
summary(model)
MuMIn::r.squaredGLMM(model)
#           R2m       R2c
#[1,] 0.2796572 0.7935903
plot(model)
```

- cluster uno: testing-based CAN, DEU, GBR, SWE, USA
- cluster due: restrinctions(strong)-testing(mild)-based IRL, IRN, ITA, NLD, TUR
- cluster tre: mild levels of all 3 : AUT BEL CHE DNK ESP FIN FRA NOR RUS
- cluster quattro: testing and tracing based KOR and SGP
- cluster cinque:  restriction and (mild) tracing, GRC PRT ROU SAU

```{r}
heatmap(as.matrix(vcov(model)))
```

```{r}
#plot_model(model, type = "pred", terms = c("Clusters"))
plot_model(model)
```


```{r}
p <- plot_model(model, type = "re", facet.grid=FALSE) 

p[[1]]
p[[2]]
```
```{r}
plot_model(model, type = "diag")
```

```{r}
plot_model(model, type = "pred", terms = var_LD[1])
```

```{r}
plot_model(model, type = "pred", terms = c(var_FIX[1], var_LD[1]))
```

```{r}
ranef(model)
```

# Interpretation (if we have Clusters as Fixed Effects)

```{r}
#comparison <- c("testing_based - restrinctionsStrong_testingMild_based = 0",
#                "testing_based - mild_levels_all = 0",
#                "testing_based - testing_tracing_based = 0",
#                "testing_based - restriction_tracingMild = 0",
#                "restrinctionsStrong_testingMild_based - mild_levels_all = 0",
#                "restrinctionsStrong_testingMild_based - testing_tracing_based = 0",
#                "restrinctionsStrong_testingMild_based - restriction_tracingMild = 0",
#                "mild_levels_all - testing_tracing_based = 0",
#                "mild_levels_all - restriction_tracingMild = 0",
#                "testing_tracing_based - restriction_tracingMild = 0")

#test<-multcomp::glht(model, linfct = mcp(Clusters = comparison),adjusted="holm")
#summary(test)
```

```{r}
#eff <- data.frame(effect("Clusters", model))
#eff$Group <- factor(eff$Group,levels = c("cnt","fep","scz", "bipo"))

#ggplot(eff, aes(x=Clusters, y=fit)) + 
#  geom_errorbar(aes(ymin=lower, ymax=upper), width=.1) +
#  geom_line(aes(group = 1)) + 
#  geom_point(aes(x=reorder(Clusters, upper, desc), y=fit))+
#  xlab("Clusters") + ylab("R0")


```

```{r}
#world <- ne_countries(scale = "medium", returnclass = "sf")
#world$id <- world$iso_a3

#dat_shape$R0fit <- ifelse(dat_shape$Clusters == "testing_based", eff$fit[1],
#                    ifelse(dat_shape$Clusters == #"restrinctionsStrong_testingMild_based", eff$fit[2],
#                           ifelse(dat_shape$Clusters == "mild_levels_all", #eff$fit[3],
#                                  ifelse(dat_shape$Clusters == #"testing_tracing_based", eff$fit[4],eff$fit[5]))))

#datMap <- left_join(world, dat_shape[,c("id","R0fit")], by="id")
#datMap$R0fit <- as.factor(datMap$R0fit)
#datMap$R0fit <- recode_factor(datMap$R0fit, "2.1226864" = "testing_based_2_12",
#                              "2.7178730" = #"restrinctionsStrong_testingMild_based_2_72",
#                              "2.3098266" = "mild_levels_all_2_31",
#                              "0.9163245" = "testing_tracing_based_0_92",
#                              "1.2222473" = "restriction_tracingMild_1_22")
#ggplot(data = datMap) + geom_sf(aes(fill = R0fit)) + scale_color_ordinal()

```

# Interpretation (considering Clusters as Random Effects).

Represent the R0_lag in three epoches to understand how change it considering the clusters

Epoche da 14 a 103, analyzing the following plot:

```{r}
maxE <- max(as.numeric(dat_shape$date2))
minE <- min(as.numeric(dat_shape$date2)[dat_shape$R0_lag>0],na.rm = TRUE)

a<-dat_shape %>% ggplot(aes(x=date2, y=R0_lag, group=Clusters, color=Clusters)) + geom_smooth() + geom_vline(xintercept = 34)

ggplotly(a)
```

We take from 14 to 34

```{r}
dat_shape$EPOCH <- ifelse(dat_shape$date2 <=34, "First", "Second")
```

Plot considering the mean during first epoch:


```{r}
world <- ne_countries(scale = "medium", returnclass = "sf") 
world <- world %>% filter(id %in% states_to_sel)
world$id <- world$iso_a3 

datMean <- dat_shape %>% filter(EPOCH == "First") %>% 
  group_by(id) %>%
  summarise(R0mean = mean(R0_lag,na.rm = "TRUE"), 
            Clusters = unique(Clusters),
            lat = unique(lat),
            lng = unique(lng))

datMap <- left_join(world, datMean, by="id")

p <-ggplot(data = datMap) + geom_sf(aes(fill = R0mean))+
  scale_fill_viridis_c(option = "C")

p + annotate("text", x = lat, y = lng, label = paste0(Clusters))

#NA??
```


